---
title: "STA 5167 Final Project"
author: "Brianna Alston and Lucas Nolting"
date: "2023-03-20"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading in the Dataset

```{r readData, echo=FALSE}
library(readxl)
#STA_5167_Heart_Disease_Data <- read_excel("C:/Users/Lucas Nolting/Downloads/STA 5167 Heart Disease Data.xlsx")
STA_5167_Heart_Disease_Data <- read.csv("C:/Users/bri4e/OneDrive/FSU/Spring_2023/sta5167/final/STA5167-Final/heart.csv")
knitr::kable(head(STA_5167_Heart_Disease_Data))
```

## Selecting Variables for the Model

### Categorical Variables

- sex (1=male; 0=female)
- cp: chest pain type
- exang: exercise induced angina (1=yes; 0=no)
- fbs: fasting blood sugar >120 mg/dl (1=true; 0=false)
- ca: number of major vessels colored by flourosopy (0-3) 
- target: heart disease (1=yes; 0=no)

### Continuous Variables

- age in years
- trestbps: resting blood pressure
- oldpeak: ST depression 
- thalach: maximum heart rate
- chol: cholesterol


```{r variableSelection, echo=FALSE}
sex      <- as.factor(STA_5167_Heart_Disease_Data$sex)
exang    <- as.factor(STA_5167_Heart_Disease_Data$exang)
fbs      <- as.factor(STA_5167_Heart_Disease_Data$fbs)
target   <- as.factor(STA_5167_Heart_Disease_Data$target)
ca       <- as.factor(STA_5167_Heart_Disease_Data$ca)
age      <- STA_5167_Heart_Disease_Data$age
trestbps <- STA_5167_Heart_Disease_Data$trestbps
oldpeak  <- STA_5167_Heart_Disease_Data$oldpeak
thalach  <- STA_5167_Heart_Disease_Data$thalach
chol     <- STA_5167_Heart_Disease_Data$chol

variables <- data.frame(age, sex, trestbps, thalach, exang,
                      oldpeak, fbs, ca, chol, target)
```

# Model Selection

## Fitting the full model

```{r fitModel}
summary(glm(target~age+sex+trestbps+thalach+exang+oldpeak+fbs+ca+chol, data = variables, family=binomial))
```
### Loading Packages

```{r loadPackages, message=FALSE,warning=FALSE}
# Packages for ROC
library(caTools)
library(ROCR) 
```

## Evaluating Model Performance

We will split the data into training and testing data sets, and evaluate the percentage of correct classifications.

### Creating the Training and Testing Data 

```{r splitData}
set.seed(12345)
split <- sample.split(variables, SplitRatio = 0.8)
training_data <- subset(variables, split == "TRUE")
test_data <- subset(variables, split == "FALSE")
```

### Fitting the model on the training data

```{r fitTraining,results='hide'}
logistic_model<-glm(target~sex+trestbps+thalach+exang+oldpeak+ca+chol, data = training_data, family="binomial")
summary(logistic_model)
```
### Predicting the likelihood of heart disease on the testing data using a cutoff value of .5

```{r predictTest}
# Predict test data based on model
predict_reg <- predict(logistic_model, 
                       test_data, type = "response")

# Changing probabilities at cutoff value
predict_reg <- ifelse(predict_reg >0.5, 1, 0)
```

### Creating a Confusion Matrix

```{r confusionMatrix,echo=FALSE}
# using confusion matrix
knitr::kable(table(test_data$target, predict_reg))
```

### Evaluating Model Accuracy

```{r modelAccuracy,echo=FALSE}
missing_classerr <- mean(predict_reg != test_data$target)
print(paste('Accuracy =', 1 - missing_classerr))
```

### Plotting the ROC Curve and Calculating the AUC

```{r plotROC}
# ROC-AUC Curve
ROCPred <- prediction(predict_reg, test_data$target) 
ROCPer <- performance(ROCPred, measure = "tpr", 
                             x.measure = "fpr")
   
auc <- performance(ROCPred, measure = "auc")
auc <- auc@y.values[[1]]

# Plotting curve
plot(ROCPer,main="ROC Curve")
abline(a = 0, b = 1)
auc <- round(auc, 4)
legend(.6, .4, auc, title = "AUC", cex = 1)
```







